---
title: 'DS 2020 - Summer 2025 - Homework #5'
author: "Your Name"
output: html_document
---

---
title: 'DS 2020 - Summer 2025 - Homework #5'
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab: Scraping (into) the Hall of Fame

## Overview

In this lab, you will:

1.  Scrape the 2025 Hall of Fame voting results
2.  Clean and reformat the data
3.  Append your results to the existing `HallOfFame` table from the `Lahman` package
4.  Save the updated dataset as a `.csv` file

All work should be done in this `.Rmd` file. Submit the `.Rmd` and your knitted `.html` (or `.docx`) to Canvas.

## Step 1: Load the page and extract the table

Use the `rvest` package to read the table of Hall of Fame voting results for 2025 from:

<https://www.baseball-reference.com/awards/hof_2025.shtml>

> Scrape the table and store it in a data frame called `hall2025`.

```{r}
# Your code here
```

## Step 2: Clean the table

Perform the following steps:

-   If needed, extract column names from the first row

-   Use `parse_number()` to convert `%`, vote counts, and ranks to numeric

-   Remove any characters (like `%` or `th`) with `gsub()` or `parse_number()`

-   Create a cleaned data frame that contains the **following 9 variables**, matching the Lahman `HallOfFame` table:

     -   `playerID`: set to `NA` for now unless you can match player names manually
     -   `yearID`: set to `2025`
     -   `votedBy`: set to `"BBWAA"`
     -   `ballots`: total number of ballots (should be same for all rows)
     -   `needed`: number of votes needed for induction
     -   `votes`: number of votes the player received
     -   `inducted`: `"Y"` for inducted, `"N"` otherwise
     -   `category`: set to `"Player"`
     -   `needed_note`: set to `NA`

You can use `head(HallOfFame)` to inspect the structure.

```{r}
# Your cleaning code here
```

## Step 3: Combine with Lahman `HallOfFame` data

Bind your cleaned table (`hall2025_clean`) to the `HallOfFame` data using `bind_rows()` or `rbind()`. Save the result to a new data frame called `final_data`.

```{r}
# Your combining code here
```

## Step 4: Save the updated dataset

Save your combined data frame to a file named `HallOfFame.csv` in your working directory.

```{r eval=FALSE}
# Your saving code here
```

## Final Notes

-   Be sure your R Markdown file knits without errors
-   Show all relevant code and intermediate outputs
-   Submit your `.Rmd` file **and** the knitted `.html` or `.docx` file to Canvas

# Homework Assignment: Data Licensing & Scraping Ethics

This homework builds on the lab by introducing data licensing and ethical considerations in web scraping.

## Part 1: Who owns the data?

1. Under what license is the R package `ggplot2` published? What does that mean for use of its built-in `diamonds` dataset?

2. Find two different versions of a "diamonds" dataset on Kaggle. For each:
   - Provide the link
   - Identify its license
   - Compare the dataset to `ggplot2::diamonds`. Are they the same or different?

```{r}
# Dataset 1:
# URL:
# License:
# Comparison code:

# Dataset 2:
# URL:
# License:
# Comparison code:
```

3. What license governs the [Iowa Liquor Sales data](https://data.iowa.gov/Sales-Distribution/Iowa-Liquor-Sales/m3tr-qhgy)? What does this allow you to do with the data?

4. Why do you think there are so many datasets on Kaggle that resemble one another without attribution?

## Part 2: Write Some Scrapers (**EXTRA CREDIT**)

If you're feeling ambitious, try any of the following:

1. Write a function `data_link_scraper()` that extracts dataset links from a Kaggle search (e.g. `https://www.kaggle.com/search?q=diamonds`)

2. Write a function `kaggle_evaluate()` that takes a dataset link and extracts metadata like license, author, and file size.

3. Write a function `same_data()` to test whether a Kaggle dataset is equivalent to `ggplot2::diamonds`.

Note: Your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission in RStudio.

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html (or Word) file with it.

(Optional but encouraged):  
If you’d like to practice using GitHub, feel free to push your `.Rmd` and knitted `.html` file to a **public GitHub repository** under your own account. If you do, paste the link to your GitHub repo below:

> GitHub repo link (optional): `__________`