---
title: 'Homework #3'
author: "Your Name"
output: html_document
---

---
title: 'Homework #3'
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Palmer Penguins

In this homework we will use the `penguins` data set. This data set is part of the `palmerpenguins` package. You can load the data set by executing the command:

```{r}
data(penguins, package="palmerpenguins")
```

If the command above throws an error message about the package not found, make sure to install the `palmerpenguins` package on your machine.

The information in the penguins data is collected by several antarctic research teams, who took body measurements of different types of penguins in three different locations. See `?penguins` for a description of all of the variables and a bit of the scientific background of this data.

1.  Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.

For all of the questions, we would like to see your answer based on R code, i.e. for each question include a code chunk with the relevant code AND write the answer to the question based on the code results in plain English.

2.  How many penguins are part of the studies? What variables were collected in the data? Show your code.

```{r}
# Your code goes here

```

**Replace this text with your answers to question 2** <br>
Rubric: 1 point (.5 for the code, 0.5 for the answer)

3.  How does the length (`bill_length_mm`) of a  penguin's beak relate to its depth (`bill_depth_mm`)?  Use the `ggplot2` package to show a chart, describe the relationship in words. Describe the structure and propose a follow-up. (3 sentences: structure of the plot, what do you see, follow-up). <br>
Rubric: 
- 1 point for constructing the plot
- 1.5 pts for the 3-sentence description
4. To investigate the group structure you see in the scatterplot above (this is a mega hint as to what you should be seeing :) ) you decide to add color to the above scatterplot. First, map island to color, then map species to color. Describe the relationship between length (`bill_length_mm`) of a  penguin's beak and its depth (`bill_depth_mm`) under these two color mappings. How do these color mappings inform the relationship between bill lengths and bill depths? <br>
Rubric: 
- 1 point for constructing the two plots
- 1.5 pts for the 3-sentence description of each plot (for a total of 3 pts)
5.  What is the exact relationship between species and islands? Draw a barchart and add some color. Make sure to describe the relationship in words.<br>
Rubric: 
- 1 point for constructing the plot
- 1.5 pts for the 3-sentence description

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission in RStudio.

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html (or Word) file with it.
