---
title: "code-2026-04-13"
author: "Ben Moolman"
date: "2026-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Web Scraping

Easy Web Scraping

```{r}
library(rvest)
url <- "https://en.wikipedia.org/wiki/2023_Baseball_Hall_of_Fame_balloting"
html <- read_html(url)
html


```

Your Turn - Reading Data

```{r}


```

Your Turn - Creating New Variables

```{r}
library(Lahman)
head(HallOfFame,2)


```

Data Munging Building a Join Key

```{r}

```

Your Turn - Identifying Problems

```{r}


```

Your Turn - Joining bbwa with People to add playerID from the People dataset

```{r}

```

SelectorGadget Result

```{r}
# Demo on all players
url <- "http://www.baseball-reference.com/players"
html <- read_html(url)
html %>% html_elements("") %>% html_text()

# Select Active Baseball Players with last name starting with A

```

Your Turn - Fernando Abad

```{r}
url <- "https://www.baseball-reference.com/players/a/abadfe01.shtml"

```

# Functions in R and your First Scraper

```{r}
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
library(rvest)
site <- read_html(url)

```

First Function

```{r}

```

Your Turn - Scraper Function

```{r}
bbwaa_scraper <- function(url) {
}
```

Your Turn - Expanding

```{r}

```

