---
title: "code-2026-03-30"
author: "Ben Moolman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

```{r}
library(tidyverse)
# install.packages("lubridate")
library(lubridate)


```

# Converter functions

```{r}

```


# Your Turn 1

Create date objects for today's date by typing the date in text format and converting it with one of the lubridate converter functions.

Try different formats of writing the date and compare the end results

```{r}


```

# Ames Liquor Sales Example

```{r}
url <- "https://github.com/DS202-at-ISU/materials/blob/master/03_tidyverse/data/ames-liquor.rds?raw=TRUE"
download.file(url, "ames-liquor.rds", mode = "wb")
ames <- readRDS("ames-liquor.rds")

```

# Accessor Functions

```{r}

```

# Intervals and Durations

```{r}

```

# Your Turn 2

Use the ames data to answer the following questions:

give a visual breakdown of the number of sales by year. Is there a pattern recognizable? Does that pattern change when considering the volume of alcohol sold?

find the weekday associated with each date in the data. Introduce a variable weekday into the dataset. What is the number of sales by weekday? What is the volume of alcohol sold? Does that pattern hold over the years?

```{r}


```

# Example Liquor Sales (cont'd) Time Series

```{r}
ames %>% 
  ggplot(aes(x = month(Date, abbr = TRUE, label=TRUE), 
             weight = `Volume Sold (Gallons)`/1000)) + 
  geom_bar() + 
  facet_wrap(~year(Date), ncol = 6) +
  ylab("Volume of alcohol sold (in thousands of  Gallons)")


```

# Your Turn 3

find the number of alcohol sales and volume sold on each day over the time frame. What is the average number of sales each day?

Why are there no 0s in the data? Does that change your assessment of the average number of sales per day?

Challenge: use the functions seq.Date and complete to introduce days into the dataset with no sales. Start by reading the help files for the functions ...

```{r}
# Hint:
# group_by(Date)
# summarise(...)
# mean(...)
# look into seq.Date() and complete()

```

