---
title: "code-2026-03-30"
author: "Ben Moolman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

```{r}
library(tidyverse)
# install.packages("lubridate")
library(lubridate)

now()
class(now())
typeof(now())

as.numeric(now())
```

# Converter functions

```{r}
x <- "2026-2-1"
class(x)
ymd(x)
class(ymd(x))

mdy("10-17-2022")
ydm("2022-17-10")
ydm("2022-35-10")
```


# Your Turn 1

Create date objects for today's date by typing the date in text format and converting it with one of the lubridate converter functions.

Try different formats of writing the date and compare the end results

```{r}
# correct cases
ymd("2026-03-30")
mdy("03-30-2026")

# incorrect cases
mdy("30-03-2026")

# shortcut for getting today's date
today()
```

# Ames Liquor Sales Example

```{r}
# load in data
url <- "https://github.com/DS202-at-ISU/materials/blob/master/03_tidyverse/data/ames-liquor.rds?raw=TRUE"
# make sure you know which working directory you are in to know where the data will download
download.file(url, "ames-liquor.rds", mode = "wb")
ames <- readRDS("ames-liquor.rds")
view(ames)
# note that each row in ames data is a sale, so total sales would be total number of rows

# look at date variable
ames$Date
class(ames$Date)
summary(ames$Date)

# Fix Date variable from character to Date type
ames$Date <- mdy(ames$Date)
# Check updated
class(ames$Date)
summary(ames$Date)
ames$Date
```

# Accessor Functions

```{r}
now()
month(now())
year(now())
wday(now(), label = TRUE)

# look at total sales for each day of the week
library(dplyr)
ames |>
  count(wday(Date, label = TRUE))
```

# Intervals and Durations

```{r}
end_date <- now()
start_date <- end_date - years(1)
start_date

# NA here since months can vary (ie February 28 or 29 days, March 31 days)
now() - months(1)
```

# Your Turn 2

Use the ames data to answer the following questions:

give a visual breakdown of the number of sales by year. Is there a pattern recognizable? Does that pattern change when considering the volume of alcohol sold?

find the weekday associated with each date in the data. Introduce a variable weekday into the dataset. What is the number of sales by weekday? What is the volume of alcohol sold? Does that pattern hold over the years?

```{r}
# Create weekday variable
ames <- ames |>
  mutate(
    weekday = wday(Date, label = TRUE)
  )
ames$weekday

# look at number of sales by year
ames |>
  ggplot(aes(x = year(Date))) +
  geom_bar()
# look at volume sold by year
ames |>
  ggplot(aes(x = year(Date),
             weight = `Volume Sold (Gallons)`)) +
  geom_bar()

# number of sales by weekday
ames %>%
  ggplot(aes(x = weekday)) +
  geom_bar()
# alternative syntax: specify data inside ggplot()
ggplot(data = ames, aes(x = weekday)) +
  geom_bar()

# total volume sold by weekday
ames |>
  ggplot(aes(x = weekday,
         weight = `Volume Sold (Gallons)`)) +
  geom_bar()
# facet by year
ames |>
  ggplot(aes(x = weekday,
         weight = `Volume Sold (Gallons)`)) +
  geom_bar() +
  facet_wrap(~year(Date))

```

# Example Liquor Sales (cont'd) Time Series

```{r}
# from class slides
ames %>% 
  ggplot(aes(x = month(Date, abbr = TRUE, label=TRUE), 
             weight = `Volume Sold (Gallons)`/1000)) + 
  geom_bar() + 
  facet_wrap(~year(Date), ncol = 6) +
  ylab("Volume of alcohol sold (in thousands of  Gallons)")
```

# Your Turn 3

find the number of alcohol sales and volume sold on each day over the time frame. What is the average number of sales each day?

Why are there no 0s in the data? Does that change your assessment of the average number of sales per day?

Challenge: use the functions seq.Date and complete to introduce days into the dataset with no sales. Start by reading the help files for the functions ...

```{r}
# Hint:
# group_by(Date)
# summarise(...)
# mean(...)
# look into seq.Date() and complete()

# compute average number of sales per day
result <- ames |>
  group_by(Date) |>
  summarise(
    number_of_sales = n(),
    volume_sold = sum(`Volume Sold (Gallons)`, na.rm = TRUE)
  )
# see variables by date
result
# get averages
mean(result$number_of_sales)
mean(result$volume_sold)

# But we have missing dates!
# If no sales were made on a specific day, then it doesn't have a row in the dataset!
# Use complete() to expand Date variable to fill in missing days between min and max of Date
# use fill = inside complete() to specify we want 0s in these places

# See how complete works without fill
result |>
  complete(
    Date = seq.Date(min(Date), max(Date), by = "day")
  )

# now add fill
result |>
  complete(
    Date = seq.Date(min(Date), max(Date), by = "day"),
    fill = list(number_of_sales = 0, volume_sold = 0)
  ) 


# repeated above piped into a summarise so it is easier to see step by step work
result |>
  complete(
    Date = seq.Date(min(Date), max(Date), by = "day"),
    fill = list(number_of_sales = 0, volume_sold = 0)
  ) |>
  summarise(
    avg_sales_per_day = mean(number_of_sales),
    avg_volume_per_day = mean(volume_sold)
  )
```

