---
title: "DS 202: Sample Exam"
date: "10/19/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# March Madness


Basketball is a multimillion dollar sports that fascinates millions of people all year round peaking in March Madness each year.
Here, we are using play-by-play data on all NBA games on two days in Oct 2017. 

## Question #1: First Overview

Read the data from [sample-combined-pbp-stats.csv](https://raw.githubusercontent.com/DS202-at-ISU/DS202-at-ISU.github.io/master/exams/sample-combined-pbp-stats.csv) without downloading the file locally. (Hint: `?read.csv`)
Each line of the file describes one play in a game. 

- How many rows and columns does the data have? What are the variables called?
- each individual game is assigned a unique identifier `game_id`. How many games are recorded in the data and how many plays (number of rows) does a game have on average?

```{r}
library(tidyverse)

# write your code here
pbp <- read.csv("https://raw.githubusercontent.com/DS202-at-ISU/DS202-at-ISU.github.io/master/exams/sample-combined-pbp-stats.csv")
dim(pbp)
str(pbp)

length(unique(pbp$game_id))

pbp %>% group_by(game_id) %>% 
  summarise(
    n = n()
  )

# recall that tally() is equivalent to summarize() then n()
# tally() will count the number of rows of each small group
pbp %>% group_by(game_id) %>% 
  tally()

# to find the average, there are multiple ways

# first way is to save your result to some variable, then 
# you can access the column you are interested and calculate
# the mean
result_1 <- pbp %>% 
  group_by(game_id) %>% 
  tally()
mean(result_1$n)

# second way is to use pull() and directly access the column 
# you are interested
pbp %>% group_by(game_id) %>% 
  tally() %>% 
  pull(n) %>% 
  mean()
```




## Question #2: Who won? 

For each game, the variable `points` keeps track of the number of points attempted in each play. 

- Use functions from the `dplyr` package to work out the number of points each team scored (check `result`). Filter out all plays that are not associated with any of the teams. 

```{r}
# after checking the dataset, we know that we should focus
# on records where `results == 'made'` in order to calculate
# the points


scores <- pbp %>% 
  filter(result == 'made') %>% 
  group_by(game_id, team) %>% # for each game and each team
  summarise(
    final_score = sum(points, na.rm = TRUE), # calculate the sum of points
    .groups = 'drop' # remove the grouping structure
  )
scores
```


- Compare your scores with the final values of home and away scores (check for `event_type` and `period`)

```{r}
pbp %>% 
  filter(event_type == 'end of period', 
         period == 4) %>% # end of period 4
  select(away_score, home_score)
```


- Using `ggplot2` plot scores by team. Sort teams by their median scores.

```{r}
# use either bar chart or scatter plot; they have the same information
scores %>% 
  ggplot(aes(x = reorder(team, final_score, median), 
             y = final_score)) + 
  geom_bar(stat = 'identity')

scores %>% 
  ggplot(aes(x = reorder(team, final_score, median), 
             y = final_score)) + 
  geom_point()
```


- In a second step, use the scores to identify winning and losing teams of each game (Hint: use dplyr again and remember what `which.max` and `which.min` are doing). 


```{r}
scores %>% 
  group_by(game_id) %>% # `mutate`(calculate or find the winning and losing team) for each game
  mutate(
    winning_team = team[which.max(final_score)],
    lossing_team = team[which.min(final_score)]
  )
```


- `player` keeps track of which player makes a shot. For each game,  identify which player in each team scored the most points. (Hint: `?rank`)

```{r}
# write your code here

# after checking the data: `player` column would be empty if it's not related 
# to scoring. Remove these rows
pbp %>% 
  filter(player != '') %>% 
  group_by(game_id, team, player) %>% # for each game and each team and each player
  summarise(
    player_total_points = sum(points, na.rm = TRUE), # find the total points
    .groups = 'drop' # although not necessary, remove the grouping structure so that we can do another group_by
  ) %>% 
  group_by(game_id, team) %>% # for each game and each team
  mutate(
    mvp = rank(-player_total_points) # calculate (not summarize) the rank based on the total points, from top to bottom
  ) %>% 
  filter(
    mvp == 1 # filter to get the first rank
  )

```



## Question #3 : On the court

The variables `converted_x` and `converted_y` give the location of the acting player on the court.  Check with the variable `event_type` to see, for which types of play we have geographic information.

For game number  `21700003` plot the geographic location of each play on the court, colour by team and incorporate visually the play's outcome (variable `result`). 

```{r}
# write your code here
```


## Question #4: Distance hurts?

Is there any evidence that shots closer to the basket are successful more often? For that,
- introduce a new variable `d` into the data that captures the distance of a player from the basket (basket is in [0,0] for variables `original_x` and `original_y`).
- draw side-by-side boxplots of distance by result and team (using ggplot2). Interpret. (Hint: you can put either `result` or `team` on the x-axis and the remaining variable into facet)

```{r}
# write your code here
```



## Question #5: Timeline

The variable `elapsed` is recorded in hour-minute-second format. Convert the information into seconds (hint: you could introduce helper variables for the conversion).
The `elapsed` time variable starts over in each period (a period has 12 mins). Introduce a new variable called `time_played` that keeps track of the time (in seconds) from the beginning of a game to the end.

Plot the timeline (`time_played`) of events (`event_type`) for game with the id `21700002` in a scatterplot. Colour by team. Make Facet by `period`. Comment on the result.

```{r}
# write your code here
```


## Question #6: Who is playing?

Variables `h1` through `h5` and `a1` through `a5` are the five players of the home team and the away team on the field at that moment in positions 1 through 5.
Reshape the data set to combine all player variables. Extract position numbers. For each game identify how many players were playing in each position. 

Draw side-by-side boxplots of the number of players by position. Comment.

```{r}
# write your code here
```



