---
title: "DS 202: Sample Exam"
date: "3/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# March Madness


Basketball is a multimillion dollar sports that fascinates millions of people all year round peaking in March Madness each year.
Here, we are using play-by-play data on all NBA games on two days in Oct 2017. 

## Question #1: First Overview

Read the data from [sample-combined-pbp-stats.csv](https://raw.githubusercontent.com/DS202-at-ISU/DS202-at-ISU.github.io/master/exams/sample-combined-pbp-stats.csv) without downloading the file locally. 
Each line of the file describes one play in a game. 

- How many rows and columns does the data have? What are the variables called?
- each individual game is assigned a unique identifier `game_id`. How many games are recorded in the data and how many plays (number of rows) does a game have on average?

```{r}
# write your code here
```




## Question #2: Who won? 

For each game, the variable `points` keeps track of the number of points attempted in each play. 

- Use functions from the `dplyr` package to work out the number of points each team scored (check `result`). Filter out all plays that are not associated with any of the teams. 

- Compare your scores with the final values of home and away scores (check for `event_type` and `period`)

- Using `ggplot2` plot scores by team. Sort teams by their median scores.

- In a second step, use the scores to identify winning and losing teams of each game (Hint: use dplyr again and remember what `which.max` and `which.min` are doing). 

- `player` keeps track of which player makes a shot. For each game,  identify which player in each team scored the most points. 

```{r}
# write your code here
```



## Question #3 : On the court

The variables `converted_x` and `converted_y` give the location of the acting player on the court.  Check with the variable `event_type` to see, for which types of play we have geographic information.

For game number  `21700003` plot the geographic location of each play on the court, colour by team and incorporate visually the play's outcome (variable `result`). 

```{r}
# write your code here
```


## Question #4: Distance hurts?

Is there any evidence that shots closer to the basket are successful more often? For that,
- introduce a new variable `d` into the data that captures the distance of a player from the basket (basket is in [0,0] for variables `original_x` and `original_y`).
- draw side-by-side boxplots of distance by result and team (using ggplot2). Interpret.

```{r}
# write your code here
```



## Question #5: Timeline

The variable `elapsed` is recorded in hour-minute-second format. Convert the information into seconds (hint: you could introduce helper variables for the conversion).
The `elapsed` time variable starts over in each period (a period has 12 mins). Introduce a new variable called `time_played` that keeps track of the time (in seconds) from the beginning of a game to the end.

Plot the timeline (`time_played`) of events (`event_type`) for game with the id `21700002` in a scatterplot. Colour by team. Comment on the result.

```{r}
# write your code here
```


## Question #6: Who is playing?

Variables `h1` through `h5` and `a1` through `a5` are the five players of the home team and the away team on the field at that moment in positions 1 through 5.
Reshape the data set to combine all player variables. Extract position numbers. For each game identify how many players were playing in each position. 

Draw side-by-side boxplots of the number of players by position. Comment.

```{r}
# write your code here
```



